Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems.

### 2023-11-28 article.qmd
Rewriting the introduction

One limitation of these tools is that they are not extensible from Python,
as they are either built on or use wrapped libraries in other languages.
This means that they allow you to "control" the model definition and simulation in a Python script,
but you cannot use other packages from the vast Python community to change the way the simulation is carried out.
For instance,
to do a Random Parameter Scan in COPASI,
we must choose between one of the predefined distributions,
but cannot use one of the meny distributions from `scipy.stats`.
In a way, we could say that
they make it easier to avoid having to export the simulation results to a CSV file
and then load them into Python for further analysis.

Another weakness is related to the way models are written.
While most tools support standards such as the System Biology Markup Language (SBML) [@huckaSBMLL3V2],
no one is expected to write an SBML file directly.
Tellurium uses a Domain Specific Language (DSL) called Antimony [@smithAntimonyModularModel2009],
which can be translated to SBML.
A DSL allows code to be reused in different programming environments,
but it will not be recognized by Integrated Development Environments (IDEs)
and therefore cannot provide the development experience we have come to expect, such as
syntax highlighting, code completion, refactoring, and static analysis.
This has been recognized by the community,
and an extension for Visual Studio Code has been created [@maTestingDynamicalSystems2021].
However, maintaining it in the future could be burdensome for the smaller systems biology community.
TODO: this is missing for other tools

>>>>

One limitation of these tools is that they are not extensible from Python,
as they are not fully Python packages
but wrap libraries in other languages to do the simulation.
As such, they allow you to "control" the model definition and simulation in a Python script.
But it is not possible to use other packages from the vast Python community to change the way the simulation is carried out.
For instance,
to do a Random Parameter Scan in COPASI,
we must choose between one of 4 predefined distributions,
but cannot use one of the many distributions from `scipy.stats`.

Another weakness is related to the way models are written.
While most tools support standards such as the System Biology Markup Language (SBML) [@huckaSBMLL3V2],
no one is expected to write an SBML file directly.
Tellurium uses a Domain Specific Language (DSL) called Antimony [@smithAntimonyModularModel2009],
which can be translated to SBML.
A DSL allows code to be reused in different programming environments,
but it will not be recognized by Integrated Development Environments (IDEs)
and therefore cannot provide the development experience we have come to expect, such as
syntax highlighting, code completion, refactoring, and static analysis.
This has been recognized by the community,
and an extension for Visual Studio Code has been created [@maTestingDynamicalSystems2021].
However, maintaining it in the future could be burdensome for the smaller systems biology community.
